Union and intersection contracts are hard, actually

### 2021-06-09 paper.tex
Clean up

Keil and Thiemann realize the tour de force of combining union and intersection
contracts with user-defined contracts. The price to pay, though, is that
implementing their solution in a reasonable way is far from obvious. Not only
when starting from the more abstract, non-deterministic formulation, but even
when starting from the algorithmic version of their system.

% Another difficult issue Keil and Thiemann were able to solve, is the
% fact that user defined contracts and intersection do not usually play nice
% together, as shown on Section \ref{sec:flat-and-inter}.
% They provide two different sets of rules for their calculus,
% the first one is non deterministic, and its purpose is to provide
% semantics that are simple to understand.
% With these rules, the interaction between user defined contracts and
% intersections becomes much simpler, since they are able to check
% both sides of an intersection on separate threads.
% This allows them to be sure that contracts from one branch of the intersection
% won't get blamed while executing the other branch, and vice versa.
%
% As an example, consider the programs shown on Figure
% \ref{fig:kt-non-deterministic}, if the Original program
% gets executed, it will spawn two instances of said program,
% one with the first branch of the intersection (\nickel{3 | Num}),
% and the other one with the second branch (\nickel{3 | Positive}),
% and these will get joined once they reach an equal point, in this case
% it will be \nickel{3}.
%
% \begin{figure}[h]
% \begin{lstlisting}[language=nickel, title=Original]
% 3 | Num @& Positive
% \end{lstlisting}
% \begin{lstlisting}[language=nickel, title=First Program]
% 3 | Num
% \end{lstlisting}
% \begin{lstlisting}[language=nickel, title=Second Program]
% 3 | Positive
% \end{lstlisting}
% \begin{lstlisting}[language=nickel, title=Joined Program]
% 3
% \end{lstlisting}
% \caption{Non deterministic intersection contract opening}
% \label{fig:kt-non-deterministic}
% \end{figure}
%
% The second set of rules is deterministic, and it is intended
% to be simpler to implement.
% However, this second set needs to perform operations
% on the evaluation context, to decide whether or not to evaluate
% a given contract.
% \footnote{You can check more on this topic on Section 5.1 of
% \cite{KeilThiemannUnionIntersection}}
% For instance, in the program presented on Figure \ref{fig:inter-flat-contracts},
% when using the value \nickel{(fun x => x) | Str -> Str)}
% inside the execution of contract \nickel{C}, and applying it to \nickel{0},
% the contract runtime would check the context of execution and it would
% deduce that contract \nickel{Str -> Str} does not apply in this case,
% skipping it.
%
% We see two problems here, a minor one is the fact that the two
% different semantics defined for the calculus are really far apart
% one from the other, both on the strategies used for contract resolution,
% as well as on the complexity levels.
\resolved{(Yann) Is it really a
    problem per se? It is common thing to do (a nice declarative system
    for proofs and an algorithmic one for implementation, proved equivalent).
    Maybe the point is that the algorithmic system is way more complex than the
declarative one
(Teo) reworded}.
% A second, major one,
% is the dependency added over the evaluation context, since, again,
% it makes composing this, with other features desired on programming languages,
% difficult, or even impossible.
\resolved{(Yann) I think we need to substantiate this claim. If this
context dependency explains why CSE is invalid, maybe have a little example?
(Teo) I removed it, I don't think there's an example, mainly since
this context dependency is not part of the language. I do think it complicates
implementation of said feature}
\todo{(Yann) TODO: see if there's a low hanging complexity
    argument we can make (like it's quadratic in the number of open contracts or
whatever)}

>>>>

Keil and Thiemann realize the tour de force of combining union and intersection
contracts with user-defined contracts. The price to pay, though, is that
implementing their solution in a reasonable way is far from obvious. Not only
when starting from the more abstract, non-deterministic formulation, but even
when starting from the algorithmic version of their system.
\todo{(Yann) TODO: see if there's a low hanging complexity
    argument we can make (like it's quadratic in the number of open contracts or
whatever)}
